Improving the reliability of cognitive task measures: \break A narrative review

### 2022-10-28 manuscript/main.tex
latest (final?) draft

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}

@@

\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena. However, many recent studies have found that task measures exhibit poor reliability, hampering their utility for individual differences research. The purpose of this article is to provide a narrative review of approaches to improve the reliability of cognitive task measures. First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks. Then, we introduce a taxonomy of approaches for improving task reliability. Where appropriate, we highlight studies that are exemplary for improving the reliability of specific task measures. We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual-differences research.}

@@

Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry. Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in our everyday lives (e.g. reward sensitivity in a monetary-incentivized task and anhedonia in depression). Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{spiegel2021relations} and cognitive decline \cite{hartshorne2015does}. Cognitive tasks, then, are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes. For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/fig01.png}
    \caption{\small The relationship between measure reliability, observed correlations, and statistical power. (A) The maximum expected observed correlation between two measures as a function of their true (latent) correlation and reliability. As the reliability of two measures decreases, so too does their observed correlation. (B) Required sample size for 80\% power to detect true correlations between two measures given their reliability. As the reliability of two measures decreases, the number of participants required to detect an association increases. Even with a large sample size of $N=500$, two measures with moderate reliability ($\rho_{xx'} = \rho_{yy'} = 0.5$) will only reliably detect true correlations above $0.3$, which are ostensibly high for the field.}
    \label{fig:fig01}
\end{figure}

@@

Although verifying the reliability of cognitive task measures is paramount to individual differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}. When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered the minimum acceptable level for individual-differences research ($\rho_{tt}$ on the order of 0.7 -- 0.8) and the reliability regularly achieved by self-report measures. Indeed, many studies have now found that task measures exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.

@@

There are two measures of reliability of relevance for task measures: internal consistency and test-retest reliability. Internal consistency is the reliability of a measure in a single administration of a task. Test-retest reliability (also known as temporal stability) is the reliability of a task measure across two or more administrations. Even for cross-sectional experiments, where stability over time is of no scientific interest, test-retest reliability is a useful index. This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant, state-dependent factors (e.g. current mood, fatigue). Given sufficient time between testing sessions, test-retest reliability should be less biased by state factors, and thus a better estimate of the true cross-sectional task-measure reliability. However, whereas high internal consistency is always desirable for individual difference studies, high test-retest reliability may not be, depending on the construct a researcher is intending to measure. Low one-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g., mood-dependent attentional biases), but is a problem if individual differences in a cognitive process are hypothesized to be stable across time (e.g., extraversion and social reward processing).

The test-retest reliability of a task measure can be calculated in numerous ways \cite{Parsons2019-jw}. Perhaps the simplest approach is to compute the Pearson correlation between participants' scores from two sessions. An alternative approach is to calculate the intraclass correlation coefficient (ICC), which decomposes a task measure into true score variance and error variance. There are many formulas for calculating ICC \cite{Koo2016-fz}, with the critical distinction being whether reliability is based on the \textit{consistency} or \textit{absolutely agreement} of a task measure across two administrations. Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g., due to practice effects on task performance). In contrast, absolute-agreement based ICCs measure the degree to which scores are stable across time. The type of ICC to use depends on the experimenter's goals and the ultimate use of the task measure. % desiderata from the measured individual differences. %If an experimenter cares only about the rank-ordering of participants across time, a consistency-based ICC (or Pearson correlation) is preferable. Conversely, if systematic differences in a task measure across task administrations is a concern, then an absolute agreement-based ICC should be preferred.

Calculating the internal consistency of a task measure is more complicated. The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for many task data (e.g., equivalence of trials, uncorrelated measurement error). As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability, where $K$ indicates the number of splits. Split-half reliability ($K=2$) is most common. A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}. For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions due to practice, fatigue, or other linear time effects. In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may be confounded when behavior across trials is non-independent (i.e., measurement error is correlated across trials). Under such conditions, odd-even splitting will artificially inflate the similarity of data across partitions, thereby decreasing estimates of measurement noise and overestimating reliability. Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended \cite{pronk2022can, Parsons2019-jw}. Here, reliability is averaged across many thousands of random partitions of the data into halves. For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the task measure for each new subset of the data. Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive model-based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials. In these cases, one can design tasks with at least two independent blocks. Then a model is fit to each block independently and reliability is calculated using the model parameters estimated from each.

As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning}. These summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increase the measurement error in (and thus diminish the reliability of) these measures of task performance. Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior. By directly modeling trial-by-trial variability, estimates of participants' performance are effectively de-noised. It is also advisable to use hierarchical models that partially pool data across participants, which in turn further increases the precision of estimates of participants' performance. Together, these methods can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}, as multiple studies have demonstrated  \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}. For a detailed discussion of generative models in the context of task reliability, see Haines (this issue).

@@

One strategy for minimizing practice effects is simply to increase the time interval between task administrations. The more time that elapses between sessions, the greater the probability that participants will have forgotten task-specific knowledge or strategies \cite{hausknecht2007retesting, scharfen2018retest}. Of course, this solution may not always be possible or desirable, especially if a researcher is only able to or specifically interested in studying a behavior over a short time period. Moreover, some forms of learning do not easily dissipate with time.

A second strategy is to design tasks so as to prevent or discourage the formation of task-specific strategies. For example, McLean and colleagues \cite{mclean2018towards} investigated the repeatability of the beads task. In the beads task, participants are presented with two jars containing beads of two colors in equal but opposite ratios. In each trial of the task, a predetermined sequence of beads is drawn from one jar. Participants must decide which jar beads were being drawn from or request to see more beads. In a typical version of the task, the same sequence of bead draws is used across all trials. McLean and colleagues found that participants were aware that the sequence repeated across trials and, as a consequence, became more erratic in their decision to witness more bead draws with additional trials. In response, the authors developed a new version of the task that included distractor sequences of bead draws. This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores.

@@

% Similarly, Snijder and colleagues \cite{snijder2022psychometric} studied the reliability of cognitive conflict effects in multiple redesigned cognitive-control tasks (e.g., Stroop, AX-CPT, Cued Task Switching, Sternberg) that required of participants either more proactive control (i.e., maintained vigilance) or more reactive control than the traditional versions of these tasks. The authors found that the proactive control versions of the task were more demanding, which similarly led to larger cognitive interference effects and better task reliability.

A related strategy is to calibrate the difficulty of the task to the average ability of the population of interest. For example, consider a task trial with two response options. Given the Bernoulli distribution, the variance in responses on this trial will be maximal when the probability of choosing either response is equal. Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items is matched to the average ability of the sample \cite{gulliksen1945relation, ferrando2007external} (or slightly higher if participants can guess the correct response \cite{lord1952relation, feldt1993relationship}). Of course, this design principle is only helpful to the degree that a researcher knows the average ability level of their participants. If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a uniform range of difficulty.

@@

Perhaps the most straightforward approach to decreasing measurement error, and thereby increasing reliability, is to increase the number of task trials. In generalizability theory, the relationship between reliability and the number of trials defined as:
\begin{equation}
    \rho = \frac{\sigma^2_T}{\sigma^2_T + \frac{\sigma^2_E}{n}}
\end{equation}
where $\sigma^2_T$ is the true between-participants variance, $\sigma^2_E$ is measurement error (i.e. trial-level variance), and $n$ is the number of trials. In practice, this relationship often holds (e.g. \cite{paap2016role, cooper2017role}) with some exceptions \cite{price2015empirical, klingelhoefer2022robust}. Notably, increasing the number of task trials only benefits reliability if measurement error is random. If increasing task length results in participant fatigue or boredom, measurement noise may systematically increase and reliability will suffer. Increasing the number of trials may be impractical for other reasons, and because of diminishing marginal improvements for reliability, achieving a desired level of reliability through this means alone may require prohibitively long experiments.

@@

It is worth stressing that clear instructions are essential for task reliability. When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more variable across time (as their understanding of the task evolves) and across participants (due to different interpretation of instructions), thereby diminishing reliability. (Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using strategies not of interest to the experimenter.) In their ``10 simple rules'' paper for designing cognitive experiments \cite{barbosa2022practical}, Barbosa and colleagues provide practical suggestions for writing task instructions. When adapting a task for a new population, experimenters should ensure that the instructions are still appropriate. Task instructions that are comprehensible to healthy adult participants may not be suitable for other populations like children \cite{hughes2002measuring}.

@@

The equations above suggest three steps experimenters can take to improve the reliability of difference scores: (1) improve the reliability of the component measures, (2) increase the relative difference between the variances of two task measures, and (3) minimize the correlation between the task measures. The first strategy has been our focus so far. The second approach deserves further comment. As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference between (or the ratio of) the variances of the component measures increases. Intuitively this is because, as the variance of one (but not the other) component grows, so too does the proportion of unshared reliable variance. Figure \ref{fig:fig03}A shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the ratio of the variances (i.e. $\sigma_x / \sigma_y$) is sufficiently greater or less than 1. (Also worth noting is that component variances essentially function as weights as in a weighted average, such that difference score reliability more reflects the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to another advantage of increasing the between-participants variability of a measure of performance in one experiment condition insofar that there is not an accompanying increase in perormance variability in a second condition. This may explain why making more difficult the incongruent trials in the tasks designed by Kucina and colleagues improved the reliability of their difference score measures \cite{kucina2022solution}.

\begin{figure}[tp]
    \centering
    \includegraphics[width=1\textwidth]{figures/fig02.png}
    \caption{\small Difference score reliability as a function of the variances and reliabilities of its component measures. (A) Difference score reliabilities as the ratio of the component measures' variances ($\sigma_x / \sigma_y$) increase, but the component reliabilities are held fixed ($\rho_{xx'} = \rho_{yy'} = 0.8$). When the variances of the component measures are equal ($\sigma_x / \sigma_y = 1$), larger correlations between the measures substantially diminish the reliability of a difference score measure. When the variances are unequal ($\sigma_x / \sigma_y > 1$), even large correlations between the measures are less deleterious for reliability. (B) Difference score reliabilities as the component reliabilities change, but the ratio of the component measures' variances is fixed ($\sigma_x / \sigma_y = 2$). When the variances are unequal, the reliability of a difference score reflects more the component measure with the larger variance.}
    \label{fig:fig03}
\end{figure}

@@

In this brief article, we have reviewed issues and research regarding cognitive task-measure reliability. Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting studies that are exemplary for improving the reliability of specific task measures. We hope we have made clear the importance of calculating (and reporting) the reliability of task measures intended for use in psychiatric research. We also hope we have provided a useful guide for experimenters who wish to design a new task, or improve an existing task, to study individual differences in cognitive processing.

Before concluding, we must emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures. Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether we actually measuring the constructs we intended to measure. Similarly, task measures may be reliable but exhibit poor ecological validity \cite{Steiner2021-oq}, thus being poor proxies for cognition in real-world settings. Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables we care about (e.g. symptoms, treatment response). Finally, there are many other practical considerations (e.g. duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully to the clinic or naturalistic contexts at scale \cite{germine2021toward}.

Despite the challenges of making task-measures reliable and valid, we remain optimistic about their current and future use in biological psychiatry. We firmly believe that, with further research efforts towards developing, documenting, and sharing reliable task paradigms, our field can make great strides towards understanding, predicting, and ultimately relieving psychiatric illness.

@@

The code used to generate the figures in this manuscript are publicly available at \break \url{https://github.com/nivlab/biopsych-reliability-review}.

@@

This research was funded in part by a National Science Foundation Graduate Research Fellowship.

@@

>>>>

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1,*]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}
\affil[*]{Corresponding author {\small (zorowitz@princeton.edu)}}

@@

\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena across individuals. However, many recent studies have found that task measures exhibit poor reliability, which hampers their utility for individual-differences research. Here we provide a narrative review of approaches to improve the reliability of cognitive task measures. First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks. Then, we introduce a taxonomy of approaches for improving task reliability. Where appropriate, we highlight studies that are exemplary for improving the reliability of specific task measures. We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing one, to achieve sufficient reliability for use in individual-differences research.}

@@

Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry. Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in everyday life (e.g. reward sensitivity in a monetary-incentivized task and anhedonia in depression). Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{spiegel2021relations} and cognitive decline \cite{hartshorne2015does}. Cognitive tasks, then, are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes. For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/fig01.png}
    \caption{\small The relationship between measure reliability, observed correlations, and statistical power. (A) The maximum expected observed correlation between two measures as a function of their true (latent) correlation and reliability. As the reliability of two measures decreases, so too does their observed correlation. (B) Required sample size for 80\% power to detect true correlations between two measures given their reliability. As the reliability of two measures decreases, the number of participants required to detect an association increases. Even with a large sample size of $N=500$, two measures with moderate reliability ($\rho_{xx'} = \rho_{yy'} = 0.5$) will only reliably detect true correlations above $0.3$, which are likely high for individual-differences cognitive research \cite{owens2021recalibrating}.}
    \label{fig:fig01}
\end{figure}

@@

Although verifying the reliability of cognitive task measures is paramount to individual differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}. When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered the minimum acceptable level for individual-differences research ($\rho_{xx'}$ on the order of 0.7 -- 0.8) and the reliability regularly achieved by self-report measures. Indeed, many studies have now found that task measures exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.

@@

There are two measures of reliability of relevance for task measures: internal consistency and test-retest reliability. Internal consistency is the reliability of a measure in a single administration of a task. Test-retest reliability (also known as temporal stability) is the reliability of a task measure across two or more administrations. Even for cross-sectional experiments, where stability over time is of no scientific interest, test-retest reliability is a useful index. This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant, state-dependent factors (e.g. current mood, fatigue). Given sufficient time between testing sessions, test-retest reliability should be less biased by state factors, and thus a better estimate of the true cross-sectional task-measure reliability. However, whereas high internal consistency is always desirable for individual difference studies, high test-retest reliability may not be, depending on the construct a researcher is intending to measure. For example, low one-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g., mood-dependent attentional biases), but is a problem if individual differences in a cognitive process are hypothesized to be stable across time (e.g., extraversion and social reward processing).

The test-retest reliability of a task measure can be calculated in numerous ways \cite{Parsons2019-jw}. Perhaps the simplest approach is to compute the Pearson correlation between participants' scores from two sessions. An alternative approach is to calculate the intraclass correlation coefficient (ICC), which decomposes a task measure into true score variance and error variance. There are many formulas for calculating ICC \cite{Koo2016-fz}, with the critical distinction being whether reliability is based on the \textit{consistency} or \textit{absolutely agreement} of a task measure across two administrations. Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g., due to practice effects on task performance). In contrast, absolute-agreement-based ICCs measure the degree to which scores are stable across time. The type of ICC to use depends on the experimenter's goals and the ultimate use of the task measure.

Calculating the internal consistency of a task measure is more complicated. The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for many task data (e.g., equivalence of trials, uncorrelated measurement error). As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability, where $K$ indicates the number of splits. Split-half reliability ($K=2$) is most common. A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}. For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions resulting from practice, fatigue, or other linear time effects. In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may be confounded when behavior across trials is non-independent (i.e., measurement error is correlated across trials). Under such conditions, odd-even splitting will artificially inflate the similarity of data across partitions, thereby decreasing estimates of measurement noise and overestimating reliability. Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended \cite{pronk2022can, Parsons2019-jw}. Here, reliability is averaged across many thousands of random partitions of the data into halves. For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the model for each new subset of the data. Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive-model based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials. In these cases, one can design tasks with at least two independent blocks. The model can then be fit to each block independently and reliability calculated using the model parameters estimated from each.

As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning}. This is because such summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increases the measurement error (and thus diminishes reliability). Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior. By directly modeling trial-by-trial variability, estimates of participants' performance are effectively de-noised. It is also advisable to use hierarchical models that partially pool data across participants, which in turn further increases the precision of estimates of participants' performance. Together, these methods can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}, as multiple studies have demonstrated  \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}. For a detailed discussion of generative models in the context of task reliability, see Haines (this issue).

@@

One strategy for minimizing practice effects is simply to increase the time interval between task administrations. The more time that elapses between sessions, the greater the probability that participants will have forgotten task-specific knowledge or strategies \cite{hausknecht2007retesting, scharfen2018retest}. Of course, this solution may not always be possible or desirable, especially if a researcher is only able to or specifically interested in studying a behavior over a short time period. Moreover, some forms of learning do not easily dissipate with time \cite{schiller2010preventing}.

A second strategy is to use a combination of clear instructions and practice trials to help participants reach stable performance from the start of an experiment (see the ``Improving experiment designs'' section below). Another strategy is to design tasks so as to prevent or discourage the formation of task-specific strategies. For example, McLean and colleagues \cite{mclean2018towards} investigated the repeatability of the beads task. In the beads task, participants are presented with two jars containing beads of two colors in equal but opposite ratios. In each trial of the task, a predetermined sequence of beads is drawn from one jar. Participants must decide which jar beads were being drawn from or request to see more beads. In a typical version of the task, the same sequence of bead draws is used across all trials. McLean and colleagues found that participants were aware that the sequence repeated across trials and, as a consequence, became more erratic in their decision to witness more bead draws with additional trials. In response, the authors developed a new version of the task that included distractor sequences of bead draws. This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores.

@@

A related strategy is to calibrate the difficulty of the task to the average ability of the population of interest. For example, consider a task trial with two response options. Given the Bernoulli distribution, the variance in responses on this trial will be maximal when the probability of choosing either response is equal. Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items is matched to the average ability of the sample \cite{gulliksen1945relation} (or slightly higher if participants can guess the correct response \cite{lord1952relation, feldt1993relationship}). Of course, this design principle is only helpful to the degree that a researcher knows the average ability level of their participants. If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a uniform range of difficulty.

@@

Perhaps the most straightforward approach to decreasing measurement error, and thereby increasing reliability, is to increase the number of task trials. In generalizability theory, the relationship between reliability and the number of trials defined as:
\begin{equation}
    \rho = \frac{\sigma^2_T}{\sigma^2_T + \frac{\sigma^2_E}{n}}
\end{equation}
where $\sigma^2_T$ is the true between-participants variance, $\sigma^2_E$ is measurement error (i.e. trial-level variance), and $n$ is the number of trials. In practice, this relationship often holds \cite{paap2016role, cooper2017role} though with some exceptions \cite{price2015empirical, klingelhoefer2022robust}. Notably, increasing the number of task trials only benefits reliability if measurement error is random. If increasing task length results in participant fatigue or boredom, measurement noise may systematically increase and reliability will suffer. Increasing the number of trials may be impractical for other reasons, and because of diminishing marginal improvements for reliability, achieving a desired level of reliability through this means alone may require prohibitively long experiments.

@@

It is worth stressing that clear instructions are essential for task reliability. When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more variable across time (as their understanding of the task evolves) and across participants (due to different interpretation of instructions), thereby diminishing reliability. Clear instructions thus help to ensure that participants show stable behavior from the start of the experiment. (Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using strategies not of interest to the experimenter.) In their ``10 simple rules'' paper for designing cognitive experiments \cite{barbosa2022practical}, Barbosa and colleagues provide practical suggestions for writing task instructions. When adapting a task for a new population, experimenters should ensure that the instructions are still appropriate. Task instructions that are comprehensible to healthy adult participants may not be suitable for other populations like children \cite{hughes2002measuring}.

@@

The equations above suggest three steps experimenters can take to improve the reliability of difference scores: (1) improve the reliability of the component measures, (2) increase the relative difference between the variances of two task measures, and (3) minimize the correlation between the task measures. The first strategy has been our focus so far. The second approach deserves further comment. As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference (or the ratio) between the variances of the component measures increases. Intuitively, this is because as the variance of one (but not the other) component grows, so too does the proportion of unshared reliable variance. Figure \ref{fig:fig03}A shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the ratio of the variances (i.e. $\sigma_x / \sigma_y$) is sufficiently different from 1. (It is also worth noting that component variances essentially function as weights in determining the overall reliability, such that the difference-score reliability reflects more the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to the advantage of increasing the between-participants variability of a measure of performance in one experiment condition without increasing performance variability in a second condition. This may explain why Kuchina and colleagues improved the reliability of their difference score by making only the incongruent trials more difficult \cite{kucina2022solution}.

\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{figures/fig02.png}
    \caption{\small Difference score reliability as a function of the variances and reliabilities of its component measures. (A) Difference-score reliability as the ratio of the component measure variances ($\sigma_x / \sigma_y$) increases, with component reliabilities held fixed ($\rho_{xx'} = \rho_{yy'} = 0.8$). When component measures have equal variances ($\sigma_x / \sigma_y = 1$), large correlations between the measures substantially diminish the reliability of a difference score measure. When the variances are unequal ($\sigma_x / \sigma_y > 1$), even large correlations between the measures are less deleterious for reliability. (B) Difference score reliability as the component reliabilities change, with the ratio of the component-measure variances fixed ($\sigma_x / \sigma_y = 2$). When the variances are unequal, the reliability of a difference score reflects more the component measure with the larger variance (here, component $x$).}
    \label{fig:fig03}
\end{figure}

@@

We have briefly reviewed issues and research regarding the reliability of cognitive-task measures. Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting example studies that improved the reliability of specific task measures. We hope we have made clear the importance of calculating (and reporting) the reliability of task measures intended for use in psychiatric research. We also hope we have provided a useful guide for experimenters who wish to design a new task, or to improve an existing task, in order to study individual differences in cognitive processing.

We emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures. Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether they actually measure the constructs they are intended to measure. Similarly, task measures may be reliable but exhibit poor ecological validity \cite{Steiner2021-oq}, thus being poor proxies for cognition in real-world settings. Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables of interest (e.g. symptoms, treatment response). Finally, there are many other practical considerations (e.g. task duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully in the clinic or in naturalistic environments at scale \cite{germine2021toward}.

Despite the challenges of making task-measures reliable and valid, we are optimistic about their current and future use in biological psychiatry. We firmly believe that, with further research efforts towards developing, documenting, and sharing reliable task paradigms, our field can make great strides towards understanding, predicting, and ultimately relieving psychiatric illness.

@@

The code used to generate the figures in this manuscript is publicly available at \break \url{https://github.com/nivlab/biopsych-reliability-review}.

@@

We thank Rachel Bedder for helpful feedback on this manuscript. This project was made possible with support from the National Center for Advancing Translational Sciences (NCATS), a component of the National Institute of Health (NIH), under award number UL1TR003017a and a National Science Foundation Graduate Research Fellowship.

@@

\section{Citation diversity statement}

Recent work in several fields of science has identified a bias in citation practices such that papers from women and other minority scholars are under-cited relative to the number of such papers in the field \cite{dworkin2020extent, bertolero2020racial}. Here we sought to proactively consider choosing references that reflect the diversity of the field in thought, form of contribution, gender, race, ethnicity, and other factors. First, we obtained the predicted gender of the first and last author of each reference by using databases that store the probability of a first name being carried by a woman \cite{dworkin2020extent}. By this measure (and excluding self-citations to the first and last authors of our current paper), our references contain 8.92\% woman(first)/woman(last), 19.37\% man/woman, 21.92\% woman/man, and 49.79\% man/man. This method is limited in that a) names, pronouns, and social media profiles used to construct the databases may not, in every case, be indicative of gender identity and b) it cannot account for intersex, non-binary, or transgender people. Second, we obtained predicted racial/ethnic category of the first and last author of each reference by databases that store the probability of a first and last name being carried by an author of color \cite{ambekar2009name, sood2018predicting}. By this measure (and excluding self-citations), our references contain 4.81\% author of color (first)/author of color(last), 15.59\% white author/author of color, 16.92\% author of color/white author, and 62.68\% white author/white author. This method is limited in that a) names and Florida Voter Data to make the predictions may not be indicative of racial/ethnic identity, and b) it cannot account for Indigenous and mixed-race authors, or those who may face differential biases due to the ambiguous racialization or ethnicization of their names.  We look forward to future work that could help us to better understand how to support equitable practices in science.
